A novel approach to chemical tagging in lower-dimensional latent space

### 2018-12-18 article/ms.tex
words

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\package}[1]{\texttt{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}

@@

\shorttitle{Statistically consistent chemical tagging}
\shortauthors{Casey et al.}

@@

\title{A statistically consistent approach to chemical tagging}

@@

\begin{abstract}
Chemical tagging promises to distinguish unique star formation sites based on the
present day photospheric abundances of stars.  It is not known how effective this
will be.  Clustering techniques used for chemical tagging  assume statistical 
independence between abundances. Theory and observations agree that this assumption
is incorrect: some elemental abundances are produced through multiple nucleosynthetic
processes, and most nucleosynthetic processes generates multiple chemical abundances.
Here we introduce a statistically consistent approach to chemical tagging, where we
simultaneously infer the nucleosynthetic processes that produce chemical abundances,
and we perform clustering in this lower dimensional latent space. \todo{we use galah}
\todo{scales iwth data}
\end{abstract}

@@

\begin{itemize}
	\item Chemical tagging
	\item Limits on the efficacy of chemical tagging
	\item Clustering vs common traits
	\item Dimensionality of data space
\end{itemize}

@@

@@

%Factor analysis is a statistical method used to describe variability among observed, correlated variables in terms of a potentially lower number of unobserved variables called factors.

Factor analysis is a common statistical approach for describing correlated 
observations with a lower number of latent variables \citep{factor-analysis}.
\todo{Analagous approaches used in astrophysics: pca, etc}.

Similarly, clustering techniques are commonly used to identify groups in 
data space.

Here we use a combination of these methods that provides a
consistent description of the data, given the latent variables and their 
clustering in latent space. In this work the data $\vecdata$ is a 
$\NumData \times \NumDimensions$ matrix where $\NumData$ is the number of 
stars and $\NumDimensions$ is the number of chemical abundances measured 
for each star. We assume a
generative model for the data 
\begin{equation}
	\vecdata = \factorloads\factorscores + \vec{e}
	\label{eq:generative-model}
\end{equation}

\noindent{}where $\factorloads$ is a $\NumLatentFactors \times \NumDimensions$ matrix of factor
loads that is common to all data points, and the factor scores for
the $\numdata$th data point
\begin{equation}
	\factorscores_\numdata \sim \mathcal{N}(\vec\xi_\numcomponents, \vec\Omega_\numcomponents)
\end{equation}
\noindent{}are drawn from the $\numcomponents$th multivariate  normal distribution.
The factor scores for all data points $\factorscores$ is then a 
$\NumData \times \NumLatentFactors$ matrix.
 We assume $\vec{e} \sim \mathcal{N}\left(\vec{0}, \eye\specificvariance\right)$
is independent of the latent space, and $\specificvariance$ is a
diagonal matrix of $\NumDimensions$ entries. In Appendix~\ref{app:symbols}
we list all mathematical symbols used in this work and their description.
In this model each data point can be represented as being drawn
from a mixture of multivariate normal components, except the components
are \emph{clustered in the latent space} $\factorscores$ and projected
into the data space by the factor loads $\factorloads$.

@@

\todo{this model of orthogonality takes the independence into account}

Another benefit of using a mixture of common factor analysers is
for modelling stellar chemical abundances is the scaling with
computational cost. If we considered data sets of order $3\times10^7$
entries (e.g., 30 chemical abundances for $10^6$ stars) purely as a
clustering problem, then even the most efficient clustering
algorithms would incur a significant cumulative computational 
overhead by searching the parameter space for the number of
clusters, and the optimal model parameters given that number
of components. However, because the mixture of factor analyzers
approach assumes that there is a \emph{lower dimensional latent 
space} in which the data are clustered, and that clustering is 
projected into real space by common factor loads, the 
dimensionality of the clustering problem is reduced from 
$N \times D$ to $N \times J$. This reduces computational cost through
faster execution of each E-M step, and on average fewer E-M steps
needed to reach a specified convergence threshold.

@@

Here we describe the default initialisation of the model parameters, which
we have found to be robust in a large set of models. Given a number of
$\NumLatentFactors$ latent factors we initialise the factor load
$\factorloads$ entries to be drawn from a uniform distribution
$\factorloads \sim \mathcal{U}\left(-1, +1\right)$ and then ensure that
the factor loads are orthogonal such that
%TODO: Should we re-write this so that we just draw from the scipy special ortho group?
\begin{equation}
	\factorloads \factorloads\transpose = \eye \quad .
\end{equation}
%        AL = linalg.cholesky(A.T @ A)
%        A = A @ linalg.solve(AL, np.eye(self.n_latent_factors))
% 	
We then initially assign each data point as belonging to one of the
$\NumComponents$ components by drawing from a multinomial distribution
and generating the responsibility matrix $\vec\tau$. Given the initial
factor loads and assignments, we then estimate the relative weights
$\vec\pi$, the mean factor scores of each component $\scoremeans$, and
the covariance matrix of factor scores of each component $\scorecovs$.
Finally, we initialise the specific variance $\specificvariance$ in each
dimension as 1.

@@

Here we treat the generated data set as if the true number of latent factors
and the true number of components are not known. Starting with $\NumLatentFactors = 1$
and $\NumComponents = 1$, we trialled each permutation of $\NumLatentFactors$ and $\NumComponents$
until $\NumLatentFactors_{max} = 10$
and   $\NumComponents_{max} = 40$ (e.g., twice $\NumLatentFactors_{true}$ and $\NumComponents_{true}$).
For each $\NumLatentFactors,\NumComponents$ permutation we initialised the factor
loads and component assignments randomly. We performed expectation-maximization 
cycles until the relative log-likelihood improved by less than $10^{-5}$. We then
recorded the log-likelihood of the data given the model parameters, and the
Bayesian Information Criterion \citep{bic}, 
\begin{equation}
	\textrm{BIC} = Q\log{N} - 2\log\mathcal{L}\left(\data|\vec\Psi\right) \quad . \label{eq:bic}
\end{equation} 
%\noindent{}and the pseudo-Bayesian Information Criterion \citep{pseudo-bic}
%\begin{equation}
% \textrm{pseudo-BIC} = 6(1 + \gamma)\omega{}Q\log{N} - 2\log\mathcal{L}\left(\data|\vec\Psi\right) \label{eq:pseudo-bic}
%\end{equation}
%\noindent{}where $Q$ is the total number of model parameters in both equations,
%and the unknown constants $\omega$ and $\gamma$ have constraints
%$\omega \geq 1$ and $\gamma > 0$. Here we adopt $\gamma = 0.1$ and $\omega = 1$.
These metrics are shown in Figure~\ref{fig:experiment-1-gridsearch}.

@@

Unsurprisingly the log likelihood increases with increasing numbers of latent
factors $\NumLatentFactors$ and increasing numbers of components $\NumComponents$.
The lowest BIC value is found at $\NumLatentFactors = 5$
and $\NumComponents = 8$, close to the true values ($\NumLatentFactors_{true} = 5$,
$\NumComponents_{true} = 10$). The lowest pseudo-BIC value is also found at
$\NumLatentFactors = 5$, but with fewer components: $\NumComponents = 4$.

The estimated latent factor loads found during this grid search are shown in
Figure~\ref{fig:fig:experiment-1-gridsearch-factors-init-random-and-random},
where each panel shows the estimated latent factor loads found by
expectation-maximization for a given number of latent factors and components.

The latent factors $\factorloads$ in each panel are multiplied by a rotation
matrix $\mathbf{R}$ to be as similar (up to orthogonal rotation) as those found
from the same number of latent factors and $\NumComponents  =  1$.
In doing so we find that the estimated latent factors vary slightly as
the number of components $\NumComponents$ increases.

@@

\subsection{Experiment~2: Inferring astrophysically realistic latent factors}

@@

While it is trivial to find the rotation matrix $\vec{R}$ that rotates the estimated
factor loads to the true factor loads, this property of factor analyzers limits
how we can interpret the latent variables of our model. For example, if we seek
to interpret the factor loads as being something analagous to nucleosynthetic
processes, then the estimated factor loads will need to be rotated to something
that looks astrophysically plausible before we could attribute one latent factor
as being a particular nucleosynthetic process.

In this experiment we use detailed chemical abundances of metal-poor stars as
reported by \citet{Barklem:2003}, and we demonstrate the limitations in
interpreting the latent factors as nucleosynthetic processes. The data contains
14 measurements of chemical abundances for 61 stars ($N = 61$, $D = 14$): 
Al (light odd-Z), Mg and Ca ($\alpha$), Sc, Ti, Cr, Mn, Fe, Co, and Ni (Fe-peak),
as well as Sr, Y, Ba, and Eu ($s$/$r$-process). We performed a grid search for
the optimal number of latent factors $J$ and components $K$, starting from
$J = 1$ and $K = 1$, and trialling up until $J = 8$ and $K = 5$. We initialised
the model parameters randomly for each permutation of $J$ and $K$, and ran E-M
until a convergence threshold of $10^{-5}$ was reached. In Figure~\ref{fig:exp2-gridsearch-contours}
we show the negative log likelihood and BIC recorded for the optimised set of
model parameters for each permutation of $J$ and $K$. Hyperparameters of $J = 5$
and $K = 1$ are preferred by BIC for this data set.

@@

After demonstrating that there is some interpretability in the latent factors,
here we seek to experiment whether gravitationally bound stellar clusters can
be reliably identified using the model described in Section~\ref{sec:model}.

\subsection{Experiment 4: A subset of 14 chemical abundances in \Galah}
\label{sec:experiment-galah}

@@

We included the following fourteen chemical abundances: Na, Mg, 
Si, Ca, Sc, Ti, Cr, Mn, Fe, Ni, Cu, Zn, Ba, and Eu.
By requiring that each star has \texttt{flag\_cannon = 0} and
\texttt{flag\_<x> = 0} (where \texttt{<x>} is the relevant label)
for each chemical abundance, this leaves
us with a sample of 12,798 stars each with reliable stellar
parameters ($\teff$, $\logg$) and chemical abundances.
This is relevant to the approach we use here, as the method
described here requires no missing data.
These elements are produced from multiple nucleosynthetic
pathways, and it is this subsample that we will use for 
our initial chemical experiment with \Galah\ data.

%\subsubsection{Search strategy} \label{sec:experiment-galah-search}

The expectation-maximization procedure outlined in Section \ref{sec:methods} describes 
the process of refining our estimate of the model parameters $\vec\psi$ until
the improvement in log probability between iterations decreases below some
pre-defined threshold. This method is valid for a pre-defined number of latent
factors, and a pre-defined number of components to cluster in latent space.
In practice these values are not known. For the purposes of this work, we 
perform a grid-based search in $J$ and $K$, perform E-M for each combination of
$J$ and $K$, and record the optimal parameters.

We trial the number of latent factors from $J = 1$ to $J = 12$, and trial the
number of clustering components from $K = 1$ to $K = 40$. We re-initialise the
mixture for each new trial of $\{J, K\}$, and do not make use of results from
`nearby' mixtures. After the E-M threshold of $10^{-5}$ has been reached, we
record the log likelihood of the data given the model parameters, the
Bayesian Information Criterion \citep{bic}, 
\begin{equation}
	\textrm{BIC} = Q\log{N} - 2\log\mathcal{L}\left(\data|\vec\Psi\right) \label{eq:bic}
\end{equation} 
\noindent{}and the pseudo-Bayesian Information Criterion \citep{pseudo-bic}
\begin{equation}
 \textrm{pseudo-BIC} = 6(1 + \gamma)\omega{}Q\log{N} - 2\log\mathcal{L}\left(\data|\vec\Psi\right) \label{eq:pseudo-bic}
\end{equation}
\noindent{}where $Q$ is the total number of model parameters in both equations,
and the unknown constants $\omega$ and $\gamma$ have constraints
$\omega \geq 1$ and $\gamma > 0$. Here we adopt $\gamma = 0.1$ and $\omega = 1$.

\todo{Discuss grid search, inferred factor loads, cluster variances, etc.}

\begin{figure}
	\includegraphics[width=0.45\textwidth]{experiments/galah-experiment-2-gridsearch-ll-contours.png}
	\includegraphics[width=0.45\textwidth]{experiments/galah-experiment-2-gridsearch-bic-contours.png}
	\includegraphics[width=0.45\textwidth]{experiments/galah-experiment-2-gridsearch-pseudobic-contours.png}
    \caption{Model performance metrics for Experiment 2 using a subset of 14 
    		 \Galah\ chemical abundances measured from 12,798 stars. The top 
		 	 panel shows the negative log-likelihood 
			 $-\log{\mathcal{L}\left(\data|\vec\Psi\right)}$ 
			 evaluated at each combination of latent factors $J$ and number 
			 of clusters $K$, the middle panel shows the Bayesian Information
			 Criterion (Equation \ref{eq:bic}), and the lower panel shows
			 the pseudo-Bayesian Information Criterion 
			 (Equation~\ref{eq:pseudo-bic}). The white marker indicates the 
			 lowest value in each panel.}
    \label{fig:experiment-2-gridsearch}
\end{figure}

\section{Results} \label{sec:results}

@@

\section{Conclusions} \label{sec:conclusion}

>>>>

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\package}[1]{\texttt{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\article}{\emph{Article}}

@@

\shorttitle{A novel approach to chemical tagging in lower-dimensional latent space}
\shortauthors{Casey et al.}

@@

\title{A novel approach to chemical tagging in lower-dimensional latent space}

@@

\begin{abstract}
Chemical tagging promises to distinguish unique star formation sites based on the
present day photospheric abundances of stars. Clustering techniques used to date
have treated all chemical abundances as independent, which we know is not true.
Nucleosynthetic processes make multiple elements in varying quantities, and the
dimensionality of chemical abundance space is typically lower than the reported
number of abundances. Here we introduce a novel approach to chemical tagging 
where a set of latent factors (e.g., nucleosynthetic yields) contribute to the
data, and there is clustering in the latent space (e.g., relative counts of
nucleosynthetic events). This approach allows us to infer contributions akin to
nucleosynthetic yields, and the clustering becomes more efficient as it occurs in
a lower dimensional latent space.
We run experiments on literature chemical abundances,
including the second \Galah\ data release. We find that 5 latent factors are
preferred to explain the chemical abundances of metal-poor stars, and 8 are
needed to account for \Galah\ data.
\todo{something about the clusters or factors found}
We release code associated with this \article\ that scales well with the 
available data: enabling chemical tagging for $10^{6}$ stars on a standard
desktop machine in minutes.
\end{abstract}

@@

Chemical tagging describes the process of identifying individual star formation
sites based on the present day photospheric abundances of stars \citep{Freeman;Bland-Hawthorn:2001}.
Since the dynamical dissipation timescale for open clusters is relatively short
\citep[\todo{timescale}][]{}, the ability to identify star formation sites long
after the stars are no longer gravitationally bound could provide key insights
on the early formation of the disc of the Milky Way galaxy. For this reason,
much effort has been placed on obtaining observational data, methods for analysis,
theory, etc, in order to improve the prospects of chemical tagging.

@@

\todo{early approaches on clustering metrics}

\todo{work on lower limits of chemical homogeneity}

All clustering techniques applied to chemical tagging assume that the data
dimensions are independent. However, theory and observations agree this cannot
be true. Nucleosynthetic processes produce multiple elements in varying
quantities, and the dimensionality of stellar abundance datasets has been shown
to be lower than the actual number of dimensions. Any clustering approach that
treats each new elemental abundance as an indenpendent axis of information will
therefore conclude with biased inferences about the star formation history of
our Galaxy. On the other hand, it is difficult to prescribe the nucleosynthetic
yields that have contributed to the chemical abundances of each star. There are
qualitative statements that can be made for large numbers of stars, or particular
types of stars, but quantifying the precise contribution of different processes
to each star is an unsolved problem. For example, the [$\alpha$/Fe] `knee' in
abundance ratios in the Milky Way can qualitatively be explained by 
core-collapse supernovae being the predominant nucleosynthetic process in the
early Milky Way before Type Ia contributions had a significant impact, but 
efforts to date have not sought to try to explain the detailed abundances of a 
single star as a contribution of yields from different systems (e.g., 20\% of the
Ba was produced by 3 M$_\odot$ systems). This is in part because of the
challenging and degenerate nature of the problem as described, and is complicated
by the differences in yield predictions that account from prescriptions used in
different codes.

In this work we present a new approach to chemical tagging that allows us to 
identify the latent factors that contribute to the chemical abundances of all 
stars (e.g., something akin to nucleosynthetic yields), while simultaneously
performing clustering in the latent space. Notwithstanding caveats that we will
discuss in detail, this allows us to infer nucleosynthetic yields rather than
strictly prescribe them from models. Moreover, the scale of the clustering
problem reduces by a significant fraction because the clustering is performed in
a lower dimensional latent space instead of the higher dimensional data space.
In Section~\ref{sec:methods} we describe the model and the methods we use to
estimate the model parameters. Section~\ref{sec:experiments} describe the 
experiments performed using generated and real data sets. We discuss the results
of these experiments in Section~\ref{sec:discussion}, including the caveats with
the model as described. We conclude in Section~\ref{sec:conclusion}

@@

Factor analysis is a common statistical approach for describing correlated 
observations with a lower number of latent variables \citep{factor-analysis}.
Related techniques include principal component analysis \citep{pca} and its
variants \citep{ppca}, singular value decomposition \citep{svd}, and other
matrix factorization methods. While factor analysis on its own is a useful
dimensionality reduction tool to identify latent factors that contribute to
the chemical abundances of stars, it does not perform any clustering.
Similarly, clustering techniques applied to chemical abundances to date 
\citep[e.g.,][]{Hogg:2016} do not account for the lower dimensionality of the
data.

Here we use a combination of factor analysis and a gaussian mixture model to
provide a consistent description of the data, given the latent variables and 
their clustering in latent space. In this work the data $\vecdata$ is a 
$\NumData \times \NumDimensions$ matrix where $\NumData$ is the number of 
stars and $\NumDimensions$ is the number of chemical abundances measured 
for each star. We assume a generative model for the data 
\begin{equation}
	\vecdata = \factorloads\factorscores + \vec{e}
	\label{eq:generative-model}
\end{equation}

\noindent{}where $\factorloads$ is a $\NumLatentFactors \times \NumDimensions$ 
matrix of factor loads that is common to all data points, and the factor scores 
for the $\numdata$th data point
\begin{equation}
	\factorscores_\numdata \sim \mathcal{N}(\vec\xi_\numcomponents, \vec\Omega_\numcomponents)
\end{equation}
\noindent{}are drawn from the $\numcomponents$th multivariate normal distribution.
The factor scores for all data points $\factorscores$ is then a 
$\NumData \times \NumLatentFactors$ matrix, where each data point has a partial
association to the components in latent space. 
We assume $\vec{e} \sim \mathcal{N}\left(\vec{0}, \eye\specificvariance\right)$
is independent of the latent space, and $\specificvariance$ is a
diagonal matrix of $\NumDimensions$ entries. 
%In Appendix~\ref{app:symbols}
%we list all mathematical symbols used in this work and their description.
In this model each data point can be represented as being drawn
from a mixture of multivariate normal components, except the components
are \emph{clustered in the latent space} $\factorscores$ and projected
into the data space by the factor loads $\factorloads$.

@@

Including latent factors in the model description allows us to account for 
processes that affect multiple elemental abundances. In this way we are 
accounting for the fact that the data dimensions are not all independent of
each other. Another benefit is the scaling with computational cost. If we 
considered data sets of order $3\times10^7$
entries (e.g., 30 chemical abundances for $10^6$ stars) purely as a
clustering problem, then even the most efficient clustering
algorithms would incur a significant cumulative computational 
overhead by searching the parameter space for the number of
clusters, and the optimal model parameters given that number
of components. However, because the mixture of factor analyzers
approach assumes that there is a \emph{lower dimensional latent 
space} in which the data are clustered, and that clustering is 
projected into real space by common factor loads, the 
dimensionality of the clustering problem is reduced from 
$N \times D$ to $N \times J$. This reduces computational cost through
faster execution of each E-M step, and on average fewer E-M steps
needed to reach a specified convergence threshold.

@@

Here we describe the default initialisation of the model parameters, which
we have found to be robust in a large set of models. Given a number of
$\NumLatentFactors$ latent factors we initialise the factor load
$\factorloads$ entries to be drawn from a uniform distribution
$\factorloads \sim \mathcal{U}\left(-1, +1\right)$ and then ensure that
the factor loads are orthogonal such that
%TODO: Should we re-write this so that we just draw from the scipy special ortho group?
\begin{equation}
	\factorloads \factorloads\transpose = \eye \quad .
\end{equation}
%        AL = linalg.cholesky(A.T @ A)
%        A = A @ linalg.solve(AL, np.eye(self.n_latent_factors))
% 	
We then initially assign each data point as belonging to one of the
$\NumComponents$ components by drawing from a multinomial distribution
and generating the $\NumData \times \NumComponents$ responsibility matrix $\vec\tau$,
where $\sum_{k=1}^{K} \tau_{nk} = 1$ for any $n$th row. Given the initial
factor loads and assignments, we then estimate the relative weights
$\vec\pi$, the mean factor scores of each component $\scoremeans$, and
the covariance matrix of factor scores of each component $\scorecovs$.
Finally, we initialise the specific variance $\specificvariance$ in each
dimension as 1.

@@

Here we treat the generated data set as if the true number of latent factors
and the true number of components are not known. Starting with $\NumLatentFactors = 1$
and $\NumComponents = 1$, we trialled each permutation of $\NumLatentFactors$ and $\NumComponents$
until $\NumLatentFactors_{max} = 10$
and   $\NumComponents_{max} = 40$ (e.g., twice $\NumLatentFactors_{true}$ and $\NumComponents_{true}$).
For each $\NumLatentFactors,\NumComponents$ permutation we initialised the factor
loads and component assignments randomly. We performed expectation-maximization 
cycles until the relative log-likelihood improved by less than $10^{-5}$. We then
recorded the log-likelihood of the data given the model parameters, and the
Bayesian Information Criterion \citep{bic}, 
\begin{equation}
	\textrm{BIC} = Q\log{N} - 2\log\mathcal{L}\left(\data|\vec\Psi\right) \quad . \label{eq:bic}
\end{equation}

These metrics are shown in Figure~\ref{fig:experiment-1-gridsearch}.
Unsurprisingly the log likelihood increases with increasing numbers of latent
factors $\NumLatentFactors$ and increasing numbers of components $\NumComponents$.
The lowest BIC value is found at $\NumLatentFactors = 5$
and $\NumComponents = 21$, close to the true values ($\NumLatentFactors_{true} = 5$,
$\NumComponents_{true} = 20$).

%\noindent{}and the pseudo-Bayesian Information Criterion \citep{pseudo-bic}
%\begin{equation}
% \textrm{pseudo-BIC} = 6(1 + \gamma)\omega{}Q\log{N} - 2\log\mathcal{L}\left(\data|\vec\Psi\right) \label{eq:pseudo-bic}
%\end{equation}
%\noindent{}where $Q$ is the total number of model parameters in both equations,
%and the unknown constants $\omega$ and $\gamma$ have constraints
%$\omega \geq 1$ and $\gamma > 0$. Here we adopt $\gamma = 0.1$ and $\omega = 1$.

@@

@@

\subsection{Experiment~2:\\Inferring astrophysically realistic latent factors}

@@

If the true values are known then it is trivial to find a rotation matrix $\vec{R}$
that rotates the estimated factor loads to the true factor loads. However, when
the true values are not known, this property of factor analysis limits the
extent to which we can interpret the latent variables in our model. For example,
if we seek to interpret the factor loads as being something analagous to 
nucleosynthetic processes, then the estimated factor loads will need to be 
rotated to something that looks astrophysically plausible before we could 
attribute one latent factor as being a particular nucleosynthetic process. 
Although having to perform this rotation is not ideal, there is still utility in
trying to identify the latent factors because the factors can only be rotated by
some valid rotation matrix. That is to say that we will not only exactly what
we have put in: the rotated factors will simply be the closest set of factor
loads that are consistent with the data.

In this experiment we use detailed chemical abundances of metal-poor stars as
reported by \citet{Barklem:2003}, and we demonstrate the utility and limitations
in interpreting the latent factors as nucleosynthetic processes. The data 
contains 14 measurements of chemical abundances for 61 stars ($N = 61$, $D = 14$): 
Al (light odd-Z), Mg and Ca ($\alpha$), Sc, Ti, Cr, Mn, Fe, Co, and Ni (Fe-peak),
as well as Sr, Y, Ba, and Eu ($s$/$r$-process). We performed a grid search for
the optimal number of latent factors $J$ and components $K$, starting from
$J = 1$ and $K = 1$, and trialling up until $J = 8$ and $K = 5$. We initialised
the model parameters randomly for each permutation of $J$ and $K$, and ran E-M
until a convergence threshold of $10^{-5}$ was reached in the log likelihood. 
In Figure~\ref{fig:exp2-gridsearch-contours} we show the negative log likelihood 
and BIC recorded for the optimised set of model parameters for each permutation 
of $J$ and $K$. The lowest BIC value is found for $J = 5$ and $K = 1$, indicating
five latent factors are preferred, and no clustering.

@@

The previous experiment has demonstrated that the latent factors can be at least
partially interpreted within the context of astrophysical processes. In this
experiment we seek to test whether gravitationally bound stellar clusters can be
reliably identified by their chemical abundances alone, and using only the model 
described in Section~\ref{sec:model}.

\todo{description  of the open cluster data used here}

\subsection{Experiment 4: Chemical abundances in the \Galah\ survey}
\label{sec:exp4}

@@

We included the following fourteen chemical abundances: \todo{Na, Mg, 
Si, Ca, Sc, Ti, Cr, Mn, Fe, Ni, Cu, Zn, Ba, and Eu}.
By requiring that each star has \texttt{flag\_cannon = 0} and
\texttt{flag\_<x> = 0} (where \texttt{<x>} is the relevant label)
for each chemical abundance, this leaves
us with a sample of \todo{12,798} stars each with reliable stellar
parameters ($\teff$, $\logg$) and chemical abundances.
This is relevant to the approach we use here, as the method
described here requires no missing data.
These elements are produced from multiple nucleosynthetic
pathways, and it is this subsample that we will use for 
our initial chemical experiment with \Galah\ data.

\todo{we perform a grid-based search for J and K}

\subsection{Experiment 5: More chemical abundances in \Galah}
\label{sec:exp5}

\todo{which abundances, only require finite and \texttt{flag\_cannon} to be OK}

@@

\todo{
\begin{itemize}
	\item dimensionality of chemical abundance space
	\item clustering approaches do not take this into account
	\item introduced a model to simultaneously account for the clustering in
		  a lower dimensional latent space
	\item there are issues, though: this is probably  not the right model
	\item identifiability
	\item orthogonal rotation
	\item can interpret these somehow as being astrophysically motivated
	\item at low metallicities how many factors do we prefer, and what can we
		  vaguely interpret these at?
	\item when we do that to large galah data sets, what is the number of factors
	\item clustering in galah
\end{itemize}
}
\section{Conclusions} \label{sec:conclusion}
